### 2015-02-26 pre_textual/agradecimento.tex
exemplo.tex: Agradecimento extraído para pre_textual/agradecimento.tex.

% Agradecimentos

\agradecimento{
    Inserir os agradecimentos aos colaboradores
    à execução do trabalho.
}

\paginaagradecimento