### 2015-05-15 complexidade.tex
Adicionado complexidade/principais_classes.tex.

\chapter{Complexidade Computacional}

% Funções de complexidade, que satisfazem aos axiomas de Blum
\newcommand{\PhiDT}{{\mathcal{T}}} % Tempo determinístico
\newcommand{\PhiDS}{{\mathcal{S}}} % Espaço determinístico
\newcommand{\PhiNT}{{\mathcal{N\!T}}} % Tempo não-determinístico
\newcommand{\PhiNS}{{\mathcal{N\!S}}} % Espaço não-determinístico

\input{complexidade/introducao}
\input{complexidade/axiomas_blum}
\input{complexidade/classes}
\input{complexidade/teorema_uniao}
\input{complexidade/medidas}
\input{complexidade/relacoes}
\input{complexidade/principais_classes}