### 2015-09-30 oraculos.tex
Dividido arquivo oraculo.tex nas seções.

\chapter{Oráculos}

\input{oraculos/definicao}
\input{oraculos/indecidibilidade}
\input{oraculos/equivalencia}
\input{oraculos/hierarquia_indecidiveis}
\input{oraculos/reducoes}