### 2015-10-06 oraculos.tex
Unidas as seções sobre problemas indecidíveis como oráculos e equivalência entre oráculos.

\chapter{Oráculos}

\input{oraculos/introducao}
\input{oraculos/definicao}
\input{oraculos/equivalencia}
\input{oraculos/hierarquia_indecidiveis}
\input{oraculos/reducoes}
\input{oraculos/hierarquia_polinomial}