### 2015-10-14 oraculos.tex
Renomeado "hierarquia de problemas indecidíveis" para "hierarquia aritmética".

\chapter{Oráculos}

\input{oraculos/introducao}
\input{oraculos/definicao}
\input{oraculos/equivalencia}
\input{oraculos/hierarquia_aritmetica}
\input{oraculos/reducoes}
\input{oraculos/hierarquia_polinomial}